Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2016-01-19 WSSSPE3/report/wg_appendix_eng_design.tex
fixes to new eng design section and appendix

Software design principles included: abstraction; coupling and cohesion; decomposition and modularization; encapsulation and information hiding; separation of interface and implementation; sufficiency completeness \& primitiveness; and separation of concerns. Similarly, user interface design principles included: learnability; user familiarity; consistency; minimal surprise; recoverability; user guidance; and user diversity. The sustainability design principles were drawn from the Karlskrona Manifesto on Sustainability Design~\cite{Becker:2014}. The manifesto states that sustainability is systemic; multidimensional; interdisciplinary; transcends the system's purpose; applies to both a system and its wider contexts; requires action on multiple levels; requires multiple timescales; changing design to take into account long-term effects doesn't automatically imply sacrifices; system visibility is a precondition for and enabler of sustainability design.
A number of sustainable software engineering principles proposed by Tate~\cite{tate2005} were also considered including: continual refinement of product and project practices; a working product at all times; continual emphasis on design; and value defect prevention over defect detection.

>>>>

Software design principles included: abstraction; coupling and cohesion; decomposition and modularization; encapsulation and information hiding; separation of interface and implementation; sufficiency completeness \& primitiveness; and separation of concerns. Similarly, user interface design principles included: learnability; user familiarity; consistency; minimal surprise; recoverability; user guidance; and user diversity. The sustainability design principles were drawn from the Karlskrona Manifesto on Sustainability Design~\cite{Becker:2014}. The manifesto states that sustainability is systemic; multidimensional; interdisciplinary; transcends the system's purpose; applies to both a system and its wider contexts; requires action on multiple levels; requires multiple timescales; changing design to take into account long-term effects doesn't automatically imply sacrifices; system visibility is a precondition for and enabler of sustainability design.
A number of sustainable software engineering principles proposed by Tate~\cite{tate:2005} were also considered including: continual refinement of product and project practices; a working product at all times; continual emphasis on design; and value defect prevention over defect detection.

### 2016-01-19 WSSSPE3/report/wg_main_eng_design.tex
fixes to new eng design section and appendix

\todo{fill in cites in this paragraph}
Software engineering principles form the basis of methods, techniques, methodologies and tools~\cite{Vliet:2008}. However, there is often a mismatch between software engineering theory and practice particulalry in the fields of compuational science and engineering, which can lead to the development of unsustainable software~\cite{Merali:2010,hettrick:2014}. Understanding and applying software engineering principles is essential in order to create and maintain sustainable software~\cite{Becker:2016}.

>>>>

Software engineering principles form the basis of methods, techniques, methodologies and tools~\cite{Vliet:2008}. However, there is often a mismatch between software engineering theory and practice particularly in the fields of computational science and engineering, which can lead to the development of unsustainable software~\cite{Merali:2010,hettrick:2014}. Understanding and applying software engineering principles is essential in order to create and maintain sustainable software~\cite{Becker:2016}.
